Risk and Trust in Open Systems:\\ \Large Towards formalising Permission and Authority, and specifying policies for Object Capability Patterns

### 2017-09-08 shupeng_msc_report.tex
update

\makeatletter
\DeclareRobustCommand{\emp}{%
  \@nomath\em \if b\expandafter\@car\f@series\@nil
  \normalfont \else \sffamily \bfseries \fi}
\makeatother
\renewcommand{\maketitlehookd}{%
\begin{abstract}
\noindent  We revisit the concepts of \textit{permission} and \textit{authority} in the works of Miller\cite{miller2006} and \linebreak Drossopoulou et al. \cite{drossopoulou2016}, and we contribute by proposing new formal definitions of permission and authority (MayAccess and MayCall predicates in our paper). Another contribution we made is that we propose and show how the concept of domination over objects can be used to reason about OCap policies, which is novel but inspired by the work of Clarke et al. on ownership types \cite{clarke1998}. We show how our definitions above can be used to reason about object-to-object interactions, and in particular, cooperation between objects and protection of objects from unknown code. Our paper then shows how the use of our concepts of permission, authority, and domination within extended Hoare logics can formally specify well-established Object-Capability (OCap) patterns, namely the DOM Tree pattern, Caretaker pattern and Membrane pattern, in the form of \textit{OCap policies} \cite{drossopoulou2013,drossopoulou2014,drossopoulou2015,drossopoulou2015b}.  In particular, we show how we can reason about protecting the properties of nodes in the DOM Tree pattern, which is inspired by the works of Maffeis et al.\cite{maffeis2010} and Devriese et al.\cite{devriese2016}, but we do so within our methodology and the capability-safe language Pony \cite{clebsch2015}. We believe our methodology and specifications of OCap patterns are less complex than other methodologies in the literature\cite{devriese2016,swasey2017}, which will allow a programmer to reason convincingly how the use of attenuating objects in capability-safe languages can help to facilitate cooperation betweeen known and unknown code, while still preventing vulnerabilities by preserving desired properties in an open system. Lastly, we end our paper with some preliminary insights on thinking about security and risk in a non-OCap system, using the Ethereum\cite{wood2014} blockchain as a case study.\end{abstract}
}
\begin{document}
\let\sf\textsf
\input{titlepage}
\raggedbottom
\thispagestyle{empty}
\tableofcontents
\thispagestyle{empty}
\clearpage
\thispagestyle{empty}
\begin{minipage}{\textwidth}
\begin{center}
\textit{Acknowledgements}
\end{center}
\small
I am immensely grateful to my supervisor Prof. Sophia Drossopoulou, for her dedication and guidance, where many ideas included in this thesis were derived and inspired from. I will miss the countless hours we have worked together on this thesis.\\

I am also thankful for the emotional support and encouragement from my friends and family; and the Monetary Authority of Singapore for offering me a scholarship to pursue this MSc degree at Imperial College London.
\end{minipage}
\maketitle
\setcounter{page}{1}
\pagestyle{fancy} % All pages have headers and footers

@@

To put our formal definitions of permission, authority, and domination in practice and demonstrate their usefulness, 
our paper provides an illustration of three OCap patterns, DOM Tree pattern, Caretaker pattern, and Membrane pattern, which can be found respectively in \cref{sec:figdom,sec:figcaretaker,sec:figmem}. Our paper also provides the corresponding OCap policies based on our formal definitions for these three patterns. In \cref{sec:treereasoning}, we choose the Dom Tree pattern as an example to demonstrate how the DOM Tree policies help us to reason comfortably about the preservation of properties of nodes in the presence of unknown code execution. A motivation for using the DOM Tree pattern is how guest/unknown code (such as third party scripted advertising or widgets) are becoming more pervasive in the web today (where details on reasoning about them within the object-capability framework can be found in the work by Maffeis et al.\cite{maffeis2010}). Lastly, in \cref{sec:ethereum}, we offer some preliminary insights on the security properties of smart contracts on the Ethereum blockchain platform that has experienced explosive growth in recent years, and we recognise that further work has to be done in the field of smart contracts on the blockchain technology.

\subsection{Overview of Paper}
We first give an overview background history on the OCap model in \cref{sec:ocapmodel}. Readers familiar with the object-capability model and its languages are encouraged to skim through this section. In \cref{sec:specs}, we formally specify the definitions of permission, authority and domination, and we use our specifications to describe and reason how objects interact with each other. Our formal specifications are broad enough to reason in general about how two objects interact in our small specification language. That said, there are benefits to using these concepts to reason within an OCap system, which we demonstrate in \cref{sec:ocapreasoning}, because the restrictions placed by OCap languages or what we call capability-safe languages allow us to enforce protection of objects more easily. We further solidify how our methodology can be used to reason convincingly in an OCap system by applying it on a DOM Tree pattern within an OCap framework, in \cref{sec:ocappatterns}. We end the paper with some preliminary insights on Ethereum in \cref{sec:ethereum}.
%----------------------------------------------------------------------------------------
%   OCap Model
%----------------------------------------------------------------------------------------
\section{Object-Capability Model}\label{sec:ocapmodel}
\subsection{From Object to Capability}
The OCap model uses the reference graph of the objects as the access graph, and strictly requires objects to interact with each other only by sending messages on object references\cite{miller2003b}. When these references cannot be forged in the system, and when there there are no default globally accessible objects (for an object to be globally accessible, the reference to use that object has to be \textit{explicitly} granted to every object globally), then it becomes necessary that for an object to call the methods of another object, an object needs to explicitly obtain the reference of the object it is calling.

@@

%----------------------------------------------------------------------------------------
%   Formal Specifications
%----------------------------------------------------------------------------------------
\newpage
\section{Formal specifications}\label{sec:specs}
In this section we describe a small object-oriented language in \cref{sec:speclang} and define predicates that allow us to specify object-capability policies in \cref{sec:permauth}. For our language, we use the definitions of modules, classes, runtime state, interpretation, execution and arising configurations from the specification language in the appendix of \cite{drossopoulou2015b}. To express more precisely how an object can interact with another object, we give formal definitions to the ideas of permission and authority which were studied but not formally defined in \cite{miller2006}. They were later studied in several ways, e.g. in the work by Maffeis et al.\cite{maffeis2010}. Our formal definition is based on but not identical to that from  the work by Drossopoulou et al.'s\cite{drossopoulou2016}. In their papers and our paper, we generally try to capture the intuition of:
\begin{itemize}
\item permission: access right of an object to invoke a behaviour of an object
\item authority: ability to cause effects
\end{itemize}
In this work however, we contribute to the literature by taking a novel approach in formally defining both permission and authority. For permission, we have four different modes permissions specified by four MayAccess predicates. Given a module M and state \losigma,:
\begin{tabbing}
MayAccess$^{Dir,Now}$(o,o$'$): \=o has an object reference of o$'$\\
MayAccess$^{Ind,Now}$(o,o$'$): o contains a reference that points to an object that has a reference\\
\>to another object---thus forming a series of object reference(s)---\\ \> where the last object in the series has an object reference of o$'$\\
MayAccess$^{Dir,Eve}$(o,o$'$): there will be some eventual state \losigma$'$ arising from \losigma\,\\ \> where M,\losigma$'$\ \loturns\ MayAccess$^{Dir,Now}$(o,o$'$) holds\\
MayAccess$^{Ind,Eve}$(o,o$'$): there will be some eventual state \losigma$'$ arising from \losigma\,\\ \> where M,\losigma$'$\ \loturns\ MayAccess$^{Ind,Now}$(o,o$'$) holds\\
\end{tabbing}
For authority, we propose a new MayCall predicate that deviates from the MayAffect predicate proposed in Drossopoulou et al.\cite{drossopoulou2016}. While their MayAffect predicate captures the idea of a field of an object being modified, our MayCall predicate is much weaker ---we mean the ability to invoke a method of an object. Therefore implicitly, we consider method invocation as an effect in the system. Given a module M and state \losigma,:
\begin{tabbing}
MayCall(o,o$'$): \=there exists a method in o that can call a method in o$'$,\\ \> such that there will be some eventual state \losigma$'$ arising from \losigma,\\ \> where the receiver in the stack has transitioned from o in \losigma, to o$'$ in \losigma$'$\\
\end{tabbing}

We further elaborate these predicates in \cref{MayAccess} and \cref{MayCall}.
\subsection{Specification Language}\label{sec:speclang}
In our small specification language, we briefly describe the definitions of module, class, runtime state, interpretation, execution, arising configurations, and assertion. All definitions in this section are repeated from the appendix of \cite{drossopoulou2015b} for the reader's convenience. Some parts from \cite{drossopoulou2015b} have also been included in the appendix of this paper. \\

@@

Note that MayAccess$^{\_,Now}$ holds imply MayAccess$^{\_,Eve}$ holds, since an eventual state \losigma$'$ that arises from \losigma\, can refer to \losigma, and therefore, the 'now' definitions implies and are stronger than the 'eventual' definitions. Also since a series of fields \loconj{f}\ can mean a singular field f, MayAccess$^{Dir,\_}$ holds imply MayAccess$^{Ind,\_}$ holds, and therefore the 'direct' definitions are stronger than the 'indirect' definitions. \\
\begin{table}[htb]
\caption{MayAccess Relations - Observations}\label{tab:accessrelations}
\centering
\begin{tabular*}{0.5\linewidth}{c|ccc}\toprule
& \bf Now & & \bf Eventually\\
\hline
\multirow{5}{*}{\rot{\bf Indirect \enspace Direct \:}} & \multirow{2}{*}{MayAccess$^{Dir,Now}$} & \loimplies & \multirow{2}{*}{MayAccess$^{Dir, Eve}$} \\
& & \lonimpliedby &  \\
& \rot{\loimpliedby} \rot{\lonimplies}& &\rot{\loimpliedby} \rot{\lonimplies} \\
& \multirow{2}{*}{MayAccess$^{Ind,Now}$} & \loimplies & \multirow{2}{*}{MayAccess$^{Ind, Eve}$} \\
& & \lonimpliedby &\\
\end{tabular*}
\end{table}\\
We surmarise the relationships between the four flavours of MayAccess in \cref{tab:accessrelations}. Note that in an object-to-object paradigm, all four flavours of MayAccess involve object capabilities. MayAccess$^{Dir,\_}$(o,o$'$) says object o has the direct reference of o$'$, while MayAccess$^{Ind,\_}$ says object o has a path of reference(s) that lead to o$'$ (that is, o has the reference of some object x$_1$, and object x$_1$ has the reference of object x$_2$ $\dots$ and object x$_n$ has the reference of object o$'$). The 'now' and 'eventual' flavours simply state that the "direct" and "indirect" predicates either they hold at \losigma\ (now) or they can hold at some arising eventual state \losigma' from \losigma\ (eventual).

@@

\textbf{Motivation for MayCall:} The motivation for defining authority as MayCall in our paper is that we would like to describe a \textit{chain} of authority between objects, without needing objects in such a chain to necessarily modify the fields of other objects in the chain. In other words, when object A wants to use object B to modify a field in C but can only do so through object B, we want to express that we can deny A from modifying C either by denying A from calling B, or deny B from calling C. Note that object A does not need to be able to modify the fields of B to eventually modify the state of C. Hence, the MayAffect predicate used in \cite{drossopoulou2016} is too strong to decompose a long chain of authority into smaller object-to-object authorities.\\

Our MayCall(o,o$'$) predicate is not concerned whether or not the state of o$'$ is modified ---the fact that a method in o$'$ can be eventually invoked by o is good enough grounds for us to say that o has authority of o$'$. Our justification is that in the object-oriented paradigm, programmers almost always enforce security of a sensitive field in a protected object to be declared private, and define some public method in the object that can modify the concerned private field based on some condition. Hence, the question of whether an \textit{untrusted} object can modify a private field in an object \sf{o$'$} can be expressed as a question of whether the untrusted object can invoke a method of \sf{o$'$}. Given our definitions of MayCall(o,o$'$), it follows that a necessary condition for modification of a private field in o$'$ is the invocation of a method of o$'$. Proving the falsity of our MayCall(o,o$'$) predicate is therefore sufficient to prove that for a field \sf{f} in o$'$, MayAffect(o,o$'$.f) does not hold.\\
\begin{logic}
\hr\\
\emp{Lemma---[Field Modification Requires Authority]}\\
\loforall \losigma$'$,f. [\=\losigma$'$\loin Arising(M,\losigma) \loand\ \loexec{o$'$.f}{\losigma$'$}\loneq\loexec{o$'$.f}{\losigma}\loimplies \loexists o. M,\losigma\ \loturns\ MayCall(o,o$'$) ]\\
\hr
\end{logic}

@@

In this section, we formally describe the ideas of \textit{isolation} and \textit{cooperation}, and how \textit{cooperation} can either be vulnerable or protected, within the OCap system. We use the word \textit{capability} (of an object), interchangeably with reference (of an object), to represent that object reference and the right to use the object reference are indistinguishable in OCap; i.e., there is no global ambient authority to prevent an object from using another object reference. Given our 4 formal definitions of \textit{permission}, it follows that o has different kinds of permission of o$'$ depending on whether o has a direct or indirect capability of o$'$, and whether we are arguing about a current or eventual state of the system. However, if we were to use \textit{our} formal definition of \textit{authority} (causing effects where effects are method invocations), we would not be able to say that an object reference embodies authority - we elaborate more in \cref{sec:isolation and cooperation}. We use the word \textit{path} interchangeably with indirect capability or permission, to emphasise that an object o need not necessarily require the direct capability of o$'$ to invoke the behaviours of o$'$---invocation of o$'$ can be done by o through a network of paths consisting of other object capabilities.\\

@@

We show an example below using \textsc{Lemma epc} with 3 objects o, y, o$'$. Let us assume o, y, and o$'$ always exists, such that the terminating recursive predicates would return true. This example then illustrates an eventual path from o to o$'$ can only exist if one of these present 6 configurations in \losigma\ holds (where x refers to the intermediate object in \textsc{epc}):
\begin{tabbing}
1) o has the capability of y, and \=y has an eventual path to o$'$---\\
\>by y having the direct capability of o$'$\\
2) o has the capability of y, and y has an eventual path to o$'$---\\
\>by o$'$ having the direct capability of y\\
3) y has the capability of o, and y has an eventual path to o$'$---\\
\>by y having the direct capability of o$'$\\
4) y has the capability of o, and y has an eventual path to o$'$---\\
\>by o$'$ having the direct capability of y\\
5) o has the direct capability of o$'$\\
6) o$'$ has the direct capability of o\\
\end{tabbing}
Note how, these 6 configurations are direct capability configurations at \losigma, but allows us to reason about whether a potential path from o to o$'$ can exist in some arising state \losigma$'$ from \losigma.\\

\subsection{Isolation and Cooperation}\label{sec:isolation and cooperation}
In this section, we reason about isolation and cooperation using our formal definition of permission and authority. Before we define isolation and cooperation, we first propose a lemma on the relationship between permission and authority:\\

\begin{logic}
\hr\\
\emp{Lemma---[Permission(Indirect,Eventual) is necessary for Authority]}\\
M,\losigma\ \loturns\ MayCall(o,o$'$) \loimplies\ MayAccess$^{Ind,Eve}$(o,o$'$)\\
\hr
\end{logic}

@@

Given lemma \textsc{epc}, we can enforce isolation of o$'$ from o given the present path configuration of a system at \losigma, and when the closed system is eventually opened such that an unknown object has permission and authority of \textit{only} object o in some arising state \losigma', we know that the unknown code will not have permission of o$'$, since o is isolated from o$'$.\\

From the contraposition of the lemma [Permission is necessary for Authority], we have the corollary:\\

@@

Therefore, it is good to know we can guarantee the internal state and behaviour of an object o$'$ in the presence of unknown code on object o through isolation, but in practice this is not entirely useful. The quintessential question in OCap systems is whether we can allow cooperation without vulnerability. What then is cooperation?\\

We say \textit{directed} cooperation from o to o$'$ is simply that o has permission of o$'$:\\

\begin{logic}
\hr\\
\emp{Definition---[Cooperation (o $\rightarrow$ o$'$)]:}\\
M,\losigma\ \loturns\ Cooperation(o,o$'$) \loiff\ M,\losigma\ \loturns\ MayAccess$^{Ind,Eve}$(o,o$'$)\\
\hr
\end{logic}

Furthermore, we make the observation between cooperation and authority that:\\
\begin{logic}
\hr\\
\emp{Observation---[Cooperation is necessary but not sufficient for authority]}\\
\loturns\ MayAccess$^{Ind,Eve}$(o,o$'$) \lonimplies\ MayCall(o,o$'$)\\
\hr
\end{logic}

@@

The problem with cooperation is when we know o$'$ has exposed methods (or public methods), and when object o is of unknown provenance. Since we assume object o is unknown, then we must assume that given the permission of o$'$, o will contain some method that will try to invoke the behaviours of o$'$. This then leads us to our next section on vulnerability and protection in the context of cooperation.

\subsection{Vulnerability and Protection}\label{sec:vulnerability and protection}
\begin{itemize}\item We assume that object o$'$ will always have some exposed public method for the rest of this section, and is the object of interest that we want to protect.\end{itemize}
For the concepts of vulnerability and protection, our arguments are based primarily on whether the source object is trusted (known) or untrusted (unknown).\\

While we can always protect o$'$ through isolation from o (o will have no paths to o$'$), what is probably more useful is whether cooperation between o$'$ and o, is possible in a \textit{protected way}. What does that mean? Notice that if we \textit{know} or \textit{trust} the code of o, then \textbf{cooperation can actually be safe}, even if o contains methods that can eventually invoke behaviours of o$'$. Suppose we have a class KnownObj which we have knowledge of and can trust:\\

>>>>

\makeatletter
\DeclareRobustCommand{\emp}{%
  \@nomath\em \if b\expandafter\@car\f@series\@nil
  \normalfont \else \sffamily \bfseries \fi}
\makeatother
\renewcommand{\maketitlehookd}{%
\begin{abstract}
\noindent  We revisit the concepts of \textit{permission} and \textit{authority} in the works of Miller\cite{miller2006} and \linebreak Drossopoulou et al. \cite{drossopoulou2016}, and we contribute by proposing new formal definitions of permission and authority (MayAccess and MayCall predicates in our paper). Another contribution we made is that we propose and show how the concept of domination over objects can be used to reason about OCap policies, which is novel but inspired by the work of Clarke et al. on ownership types \cite{clarke1998}. We show how our definitions above can be used to reason about object-to-object interactions, and in particular, cooperation between objects and protection of objects from unknown code. Our paper then shows how the use of our concepts of permission, authority, and domination within extended Hoare logics can formally specify well-established Object-Capability (OCap) patterns, namely the DOM Tree pattern, Caretaker pattern and Membrane pattern, in the form of \textit{OCap policies} \cite{drossopoulou2013,drossopoulou2014,drossopoulou2015,drossopoulou2015b}.  In particular, we show how we can reason about protecting the properties of nodes in the DOM Tree pattern, which is inspired by the works of Maffeis et al.\cite{maffeis2010} and Devriese et al.\cite{devriese2016}, but we do so within our methodology and the capability-safe language Pony \cite{clebsch2015}. We believe our methodology and specifications of OCap patterns are less complex than other methodologies in the literature\cite{devriese2016,swasey2017}, which will allow a programmer to reason convincingly how the use of attenuating objects in capability-safe languages can help to facilitate cooperation betweeen known and unknown code, while still preventing vulnerabilities by preserving desired properties in an open system. Lastly, we end our paper with some preliminary insights on thinking about security and risk in a non-OCap system, using the Ethereum\cite{wood2014} blockchain as a case study.\end{abstract}
}
\begin{document}
\let\sf\textsf
\input{titlepage}
\raggedbottom
\thispagestyle{empty}
\tableofcontents
\thispagestyle{empty}
\clearpage
\thispagestyle{empty}
\begin{minipage}{\textwidth}
\begin{center}
\textit{Acknowledgements}
\end{center}
\small
I am immensely grateful to my supervisor Prof. Sophia Drossopoulou, for her dedication and guidance, where many ideas included in this thesis were derived and inspired from.\\
\end{minipage}
\maketitle
\setcounter{page}{1}
\pagestyle{fancy} % All pages have headers and footers

@@

To put our formal definitions of permission, authority, and domination in practice and demonstrate their usefulness, 
our paper provides an illustration of three well-known OCap patterns, DOM Tree pattern, Caretaker pattern, and Membrane pattern, which can be found respectively in \cref{sec:figdom,sec:figcaretaker,sec:figmem}. Our paper also provides the corresponding OCap policies for these three patterns in \cref{sec:dompolicies,sec:caretakerpolicies,sec:membranepolicies}.  In \cref{sec:treereasoning}, we dive deeper and choose the Dom Tree pattern as an example to demonstrate how OCap policies help us to reason comfortably about the preservation of properties of nodes in the presence of unknown code execution. A motivation for focusing on the DOM Tree pattern is how guest/unknown code (such as third party scripted advertising or widgets) are becoming more pervasive in the web today (details on reasoning about them within the object-capability framework can be found in the work by Maffeis et al.\cite{maffeis2010}). Lastly, in \cref{sec:ethereum}, we offer some preliminary insights on the security properties of smart contracts on the Ethereum blockchain platform that has experienced explosive growth in recent years.

\subsection{Overview of Paper}
We first give an overview background history on the OCap model in \cref{sec:ocapmodel}. Readers familiar with the object-capability model and its languages are encouraged to skim through this section. In \cref{sec:specs}, we formally specify the definitions of permission, authority and domination, and we use our specifications to describe and reason how objects interact with each other. Our formal specifications are broad enough to reason in general about how two objects interact in our small specification language. That said, there are benefits to using these concepts to reason within an OCap system, which we demonstrate in \cref{sec:ocapreasoning}, because the restrictions placed by OCap languages or what we call capability-safe languages allow us to enforce protection of objects more easily. We further demonstrate how our methodology can be used to provide the policy specifications for three well-known OCap patterns in \cref{sec:ocappatterns}. We end the paper with some preliminary insights on Ethereum in \cref{sec:ethereum}.
%----------------------------------------------------------------------------------------
%   OCap Model
%----------------------------------------------------------------------------------------
\section{Object-Capability Model}\label{sec:ocapmodel}
\subsection{From Object to Capability}
The OCap model uses the reference graph of the objects as the access graph, and strictly requires objects to interact with each other only by sending messages on object references\cite{miller2003b}. When these references cannot be forged in the system, and when there there are no default globally accessible objects (for an object to be globally accessible, the reference to use that object has to be \textit{explicitly} granted to every object globally), then it becomes necessary that for an object to call the methods of another object, an object needs to explicitly obtain the reference of the object it is calling.

@@

%----------------------------------------------------------------------------------------
%   Formal Specifications
%----------------------------------------------------------------------------------------
\newpage
\section{Formal specifications}\label{sec:specs}
In this section we describe a small object-oriented language in \cref{sec:speclang} and define predicates that allow us to specify object-capability policies in \cref{sec:permauth}. For our language, we use the definitions of modules, classes, runtime state, interpretation, execution and arising configurations from the specification language in the appendix of \cite{drossopoulou2015b}. To express more precisely how an object can interact with another object, we give formal definitions to the ideas of permission and authority which were studied but not formally defined in \cite{miller2006}. They were later studied in several ways, e.g. in the work by Maffeis et al.\cite{maffeis2010} and Drossopoulou et al.'s\cite{drossopoulou2016}. Our formal definition is based on but not identical to that from  \cite{drossopoulou2016}. In their papers and our paper, we generally try to capture the intuition of:
\begin{itemize}
\item permission: access right of an object to invoke a behaviour of an object
\item authority: ability to cause effects
\end{itemize}
In this work however, we contribute to the literature by taking a novel approach in formally defining both permission and authority. For permission, we have four different modes permissions specified by four MayAccess predicates. Given a module M and state \losigma,:
\begin{tabbing}
MayAccess$^{Dir,Now}$(o,o$'$): \=o has an object reference of o$'$\\
MayAccess$^{Ind,Now}$(o,o$'$): o contains a reference that points to an object that has a reference\\
\>to another object---thus forming a series of object reference(s)---\\ \> where the last object in the series has an object reference of o$'$\\
MayAccess$^{Dir,Eve}$(o,o$'$): there will be some eventual state \losigma$'$ arising from \losigma\,\\ \> where M,\losigma$'$\ \loturns\ MayAccess$^{Dir,Now}$(o,o$'$) holds\\
MayAccess$^{Ind,Eve}$(o,o$'$): there will be some eventual state \losigma$'$ arising from \losigma\,\\ \> where M,\losigma$'$\ \loturns\ MayAccess$^{Ind,Now}$(o,o$'$) holds\\
\end{tabbing}
For defining authority, we propose a new MayCall predicate that deviates from the MayAffect predicate proposed in Drossopoulou et al.\cite{drossopoulou2016}. While their MayAffect predicate captures the idea of a field of an object being modified, our MayCall predicate is much weaker ---we mean the ability to invoke a method of an object. Therefore implicitly, we consider method invocation as an effect in the system. Given a module M and state \losigma,:
\begin{tabbing}
MayCall(o,o$'$): \=there exists a method in o that can call a method in o$'$,\\ \> such that there will be some eventual state \losigma$'$ arising from \losigma,\\ \> where the receiver in the stack has transitioned from o in \losigma, to o$'$ in \losigma$'$\\
\end{tabbing}

We further elaborate these predicates in \cref{MayAccess} and \cref{MayCall}, after introducing a small object-oriented language in the next section, which we would need to define \cref{MayAccess} and \cref{MayCall}. Our small language is identical to that found in the appendix of \cite{drossopoulou2015b}.
\subsection{Specification Language}\label{sec:speclang}
In our small specification language, we briefly describe the definitions of module, class, runtime state, interpretation, execution, arising configurations, and assertion. All definitions in this section are repeated from the appendix of \cite{drossopoulou2015b} for the reader's convenience. Some parts from \cite{drossopoulou2015b} have also been included in the appendix of this paper. \\

@@

\begin{table}[htb]
\caption{MayAccess Relations - Observations}\label{tab:accessrelations}
\centering
\begin{tabular*}{0.5\linewidth}{c|ccc}\toprule
& \bf Now & & \bf Eventually\\
\hline
\multirow{5}{*}{\rot{\bf Indirect \enspace Direct \:}} & \multirow{2}{*}{MayAccess$^{Dir,Now}$} & \loimplies & \multirow{2}{*}{MayAccess$^{Dir, Eve}$} \\
& & \lonimpliedby &  \\
& \rot{\loimpliedby} \rot{\lonimplies}& &\rot{\loimpliedby} \rot{\lonimplies} \\
& \multirow{2}{*}{MayAccess$^{Ind,Now}$} & \loimplies & \multirow{2}{*}{MayAccess$^{Ind, Eve}$} \\
& & \lonimpliedby &\\
\end{tabular*}
\end{table}

We illustrate the relationships between the four flavours of MayAccess in \cref{tab:accessrelations}. Note that in an object-to-object paradigm, all four flavours of MayAccess involve object capabilities. MayAccess$^{Dir,\_}$(o,o$'$) says object o has the direct reference of o$'$, while MayAccess$^{Ind,\_}$ says object o has a path of reference(s) that lead to o$'$ (that is, o has the reference of some object x$_1$, and object x$_1$ has the reference of object x$_2$ $\dots$ and object x$_n$ has the reference of object o$'$). The 'now' and 'eventual' flavours simply state that the "direct" and "indirect" predicates either they hold at \losigma\ (now) or they can hold at some arising eventual state \losigma' from \losigma\ (eventual).\\

Note that MayAccess$^{\_,Now}$ holds imply MayAccess$^{\_,Eve}$ holds, since an eventual state \losigma$'$ that arises from \losigma\, can refer to \losigma, and therefore, the 'now' definitions implies and are stronger than the 'eventual' definitions. Also since a series of fields \loconj{f}\ can mean a singular field f, MayAccess$^{Dir,\_}$ holds imply MayAccess$^{Ind,\_}$ holds, and therefore the 'direct' definitions are stronger than the 'indirect' definitions. \\

Note that MayAccess$^{\_,Eve}$ does \textbf{not} imply MayAccess$^{\_,Now}$ holds. An analogy will be object A having the capabilities of object B and C at some state \losigma. Object B and C do not have the capabilities of each other at \losigma. At some arising state \losigma' from \losigma, it is possible that object A shares the capability of C with B through introduction. Then at some state \losigma$''$ arising from \losigma$'$ (and arising from \losigma), we know B has access to C after A's introduction. Therefore M,\losigma\ \loturns\ MayAccess$^{\_Eve}$(B,C) holds since we have a possible configuration at \losigma$''$ in which B has access to C, but M,\losigma\ \loturns\ MayAccess$^{\_Now}$(B,C) does \textbf{not} hold since no such possible configuration exists at \losigma.\\

Note also that MayAccess$^{\_,Ind}$ does \textbf{not} imply MayAccess$^{\_,Dir}$ holds. An analogy will be a configuration where object A does not have the direct capability of object C, but object A has the direct capability of object B and object B has the direct capability of object C. Therefore, Object A has a path to C (through B), but it does \textbf{not} imply A has the direct capability of C.\\

@@

\textbf{Motivation for MayCall:} The motivation for defining authority as MayCall in our paper is that we would like to describe a \textit{chain of authority} between objects, without needing objects in such a chain to necessarily modify the fields of other objects in the chain. In other words, when object A wants to use object B to modify a field in C but can only do so through object B, we want to express that we can deny A from modifying C either by denying A from calling B, or deny B from calling C. Note that object A does not need to be able to modify the fields of B to eventually modify the state of C. Hence, the MayAffect predicate used in \cite{drossopoulou2016} is too strong to decompose a long chain of authority into smaller object-to-object authorities.\\

Our MayCall(o,o$'$) predicate is not concerned whether or not the state of o$'$ is modified ---the fact that a method in o$'$ can be eventually invoked by o is good enough grounds for us to say that o has authority of o$'$. Our justification is that in the object-oriented paradigm, programmers almost always enforce security of a sensitive field in a protected object to be declared private, and define some public method in the object that can modify the concerned private field based on some condition. Hence, the question of whether an \textit{untrusted} object can modify a private field in an object \sf{o$'$} can be expressed as a question of whether the untrusted object can invoke a method of \sf{o$'$}. Given our definitions of MayCall(o,o$'$), it follows that a necessary condition for modification of a private field in o$'$ is the invocation of a method of o$'$. Proving the falsity of our MayCall(o,o$'$) predicate is therefore sufficient to prove that for a field \sf{f} in o$'$, MayAffect(o,o$'$.f) does not hold:
\begin{logic}
\hr\\
\emp{Lemma---[Field Modification Requires Authority]}\\
\loforall \losigma$'$,f. [\=\losigma$'$\loin Arising(M,\losigma) \loand\ \loexec{o$'$.f}{\losigma$'$}\loneq\loexec{o$'$.f}{\losigma}\loimplies \loexists o. M,\losigma\ \loturns\ MayCall(o,o$'$) ]\\
\\
\small\textit{Note that o can possibly refer to o$'$}\\
\hr
\end{logic}

@@

In this section, we formally describe the ideas of \textit{isolation} and \textit{cooperation}, and how \textit{cooperation} can either be vulnerable, safe, or protected, within the OCap system. We use the word \textit{capability} (of an object), interchangeably with reference (of an object), to represent that object reference and the right to use the object reference are indistinguishable in OCap; i.e., there is no global ambient authority to prevent an object from using another object reference. Given our four formal definitions of \textit{permission}, it follows that o has different kinds of permission of o$'$ depending on whether o has a direct or indirect capability of o$'$, and whether we are arguing about a current or eventual state of the system. However, if we were to use \textit{our} formal definition of \textit{authority} (causing effects where effects are method invocations), we would not be able to say that an object reference embodies authority - we elaborate more in \cref{sec:isolation and cooperation}. We use the word \textit{path} interchangeably with indirect capability or permission, to emphasise that an object o need not necessarily require the direct capability of o$'$ to invoke the behaviours of o$'$---invocation of o$'$ can be done by o through a network of paths consisting of other object capabilities.\\

@@

We show an example below using \textsc{Lemma epc} with 3 objects o, y, o$'$. Let us assume o, y, and o$'$ always exists, such that the terminating recursive predicates would return true. This example then illustrates an eventual path from o to o$'$ can only exist if one of these present 6 configurations in \losigma\ holds (where y refers to the intermediate object x in \textsc{epc}):
\begin{tabbing}
1) o has the capability of y, and \=y has an eventual path to o$'$---\\
\>by y having the direct capability of o$'$\\
2) o has the capability of y, and y has an eventual path to o$'$---\\
\>by o$'$ having the direct capability of y\\
3) y has the capability of o, and y has an eventual path to o$'$---\\
\>by y having the direct capability of o$'$\\
4) y has the capability of o, and y has an eventual path to o$'$---\\
\>by o$'$ having the direct capability of y\\
5) o has the direct capability of o$'$\\
6) o$'$ has the direct capability of o\\
\end{tabbing}
Note how, these 6 configurations are direct capability configurations at \losigma, but allows us to reason about whether a potential path from o to o$'$ can exist in some arising state \losigma$'$ from \losigma.\\

\subsection{Isolation and Cooperation}\label{sec:isolation and cooperation}
In this section, we reason about isolation and cooperation using our formal definitions of permission and authority. Before we define isolation and cooperation, we first observe the relationship between permission and authority:\\

\begin{logic}
\hr\\
\emp{Observation---[Permission(Indirect,Eventual) is necessary for Authority]}\\
M,\losigma\ \loturns\ MayCall(o,o$'$) \loimplies\ MayAccess$^{Ind,Eve}$(o,o$'$)\\
\hr
\end{logic}
This observation says that in order for an object o to invoke behaviours of o$'$, it needs to be done through a path of capabilities from o to o$'$.\\

@@

Given \textsc{lemma epc}, we can enforce isolation of o$'$ from o given the present path configuration of a closed system at \losigma, and when the closed system is eventually opened such that an unknown object has permission and authority of \textit{only} object o in some arising state \losigma', we know that the unknown code will not have permission of o$'$, since o is isolated from o$'$.\\

From the contraposition of the our observation previously that [Permission is necessary for Authority], we have the corollary:\\

@@

Therefore, it is good to know that through isolation, we can guarantee the internal state and behaviour of an object o$'$ in the presence of unknown code from object o, but in practice this is not entirely useful. The quintessential question in OCap systems is whether we can allow cooperation without vulnerability. What then is cooperation?\\

We say \textit{directed} cooperation from o to o$'$ is simply that o has indirect permission of o$'$:\\
\begin{logic}
\hr\\
\emp{Definition---[Cooperation (o $\rightarrow$ o$'$)]:}\\
M,\losigma\ \loturns\ Cooperation(o,o$'$) \loiff\ M,\losigma\ \loturns\ MayAccess$^{Ind,Eve}$(o,o$'$)\\
\hr
\end{logic}
Furthermore, we make the observation between cooperation and authority that:\\
\begin{logic}
\hr\\
\emp{Observation---[Cooperation is necessary but not sufficient for authority]}\\
\loturns\ Cooperation(o,o$'$) \lonimplies\ MayCall(o,o$'$)\\
\hr
\end{logic}

@@

The problem with cooperation is when we know o$'$ has exposed methods (or public methods), and when object o is of unknown provenance. Since we assume object o is unknown, then we must assume that given the permission of o$'$, o will contain some method that will try to invoke the behaviours of o$'$. This then leads us to our next section on vulnerability, safety, and protection in the context of cooperation.

\subsection{Vulnerability, Safety, and Protection}\label{sec:vulnerability and protection}
\begin{itemize}\item We assume that object o$'$ will always have some exposed public method for the rest of this section, and is the object of interest that we want to protect.\end{itemize}
For the concepts of vulnerability, safety, and protection, our arguments are based primarily on whether the source object is trusted (known) or untrusted (unknown).\\

While we can always protect o$'$ through isolation from o (o will have no paths to o$'$), what is probably more useful is whether cooperation between o$'$ and o, is possible in a \textit{protected way}. What does that mean? Notice that if we \textit{know} or \textit{trust} the code of o, then \textbf{cooperation can actually be safe}, even if o contains methods that can eventually invoke behaviours of o$'$. Suppose we have a class TrustedObj which we have knowledge of and can trust:\\

\begin{logic}
\hr\\
\emp{Definition---[Safe cooperation]}\\
M,\losigma\ \loturns\ SafeCooperation(o,o$'$) \loiff\ o:TrustedObj \loand\ Cooperation(o,o$'$)\\
\hr
\end{logic}
